---
title: "Assignment 4"
author: "Robert Evans"
date: "April 9, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#9.1- The Titanic Data Set
This section loads the titanic dataset. This data set is a 4-dimensional table with the class (1st, 2nd, etc.), Sex, Age, and if they survived the sinking of the Titanic. In this recordset each row represents a passenger.
```{r}
#get information on dataset variables
str(Titanic)

#convert dataset to data frame
df <- as.data.frame(Titanic)
head(df)

#convert data set to a raw format
titanic.raw <- NULL
for(i in 1:4){
  titanic.raw <- cbind(titanic.raw, rep(as.character(df[,i]), df$Freq))
}
titanic.raw <- as.data.frame(titanic.raw)
names(titanic.raw) <- names(df)[1:4]
dim(titanic.raw)

#get information on dataset variables
str(titanic.raw)

#display first few rows
head(titanic.raw)

#get counts of each variable
summary(titanic.raw)

#Read the first five lines of data from Dataset.data
readLines("data/Dataset.data", n=5)

#load dataset.data into R
titanic <- read.table("data/Dataset.data", header=F)
names(titanic) <- c("Class", "Sex", "Age", "Survived")
```

#9.2- Association Rule Mining
In this section the arules package is used to create association rules between commonalities in the titanic.raw dataset.
```{r}
#Install and load the arules package
#install.packages("arules")
library(arules)

#find association rules with the default settings
rules <- apriori(titanic.raw)

#display the rule set
rules

quality(rules) <- quality(rules)
#As a note the rhs column shows values other than survived. This is solved in the next step
inspect(rules)

#create an association rule set where only the right hand side shows if the passenger survived. This is done through the appearance paramater. Notice the rhs is set to a list of Survived=No and Survived=Yes this will keep the results to just survived.
rules <- apriori(titanic.raw,
                 parameter = list(minlen=2, supp=0.005, conf=0.8),
                 appearance = list(rhs=c("Survived=No", "Survived=Yes"),
                                   default="lhs"),
                 control = list(verbose=F))
#sort rules by lift
rules.sorted <- sort(rules, by="lift")

#display new rule set. Notice this time there are only rules where the right hand side (RHS) is set to survived.
inspect(rules.sorted)
```

#9.3- Removing Redundancy
In this section redundant rules that don't provide any additional knowledge are removed. Take for example Rule number 2. {Class=2nd,Sex=Female,Age=Child}  => {Survived=Yes} Since Rule number 1: {Class=2nd,Age=Child}  => {Survived=Yes} states that everyone survived this rule does not provide any additional information.
```{r}
#find redundant rules
subset.matrix <- is.subset(rules.sorted, rules.sorted)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- FALSE
redundant <- colSums(subset.matrix, na.rm=T) >= 1
which(redundant)

#remove redundant rules
rules.pruned <- rules.sorted[!redundant]
inspect(rules.pruned)
```

#9.4- Visualizing Association Rules
This section covers different ways to visualize the association rules. These methods are: scatter plot, ballon plot, graph and parallel coordinates plot. These are provided through the arulesViz package
```{r}
#Install and load arulesViz package
#install.packages("arulesViz")
library(arulesViz)

#create a scatter plot of rules
plot(rules)

#Create a grouped matrix of rules
plot(rules, method="grouped")

#Create a line graph of rules
plot(rules, method="graph")

#Create a graph of items
plot(rules, method="graph", control=list(type="items"))

#Create a parallel coordinates plot of the rules
plot(rules, method="paracoord", control=list(reorder=TRUE))
```




