---
title: "Extra Credit Assignment 1"
author: "Robert Evans"
date: "April 7, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Preparing R for your first project
In this section the following packages are loaded and installed. The install commands are commented out so that the kintting of the word document will run.

* plyr- Allows the use of the split-apply-combine data analysis pattern
* ggplot2- Makes visualizations eaiser
* reshape2
```{r}
#install.packages("plyr")
#install.packages("ggplot2")
#install.packages("reshape2")
library(plyr)
library(ggplot2)
library(reshape2)
```

#Importing automobile fuel effiency data into R
In this section the data is imported into R. In addition information on the variables is also imported
```{r}
#import the data- note that the data can be imported directly from a zip file if the file name is known
vehicles <- read.csv(unz("data/vehicles.csv.zip", "vehicles.csv"), stringsAsFactors = F)

#view the first few lines in the data set
head(vehicles)

#load variables information
labels <- do.call(rbind, strsplit(readLines("data/variables.txt"), " - "))

#view the first few lines of variables
head(labels)
```

