---
title: "Extra Credit Assignment 1"
author: "Robert Evans"
date: "April 7, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Preparing R for your first project
In this section the following packages are loaded and installed. The install commands are commented out so that the kintting of the word document will run.

* plyr- Allows the use of the split-apply-combine data analysis pattern
* ggplot2- Makes visualizations eaiser
* reshape2
```{r}
#install.packages("plyr")
#install.packages("ggplot2")
#install.packages("reshape2")
library(plyr)
library(ggplot2)
library(reshape2)
```

#Importing automobile fuel effiency data into R
In this section the data is imported into R. In addition information on the variables is also imported
```{r}
#import the data- note that the data can be imported directly from a zip file if the file name is known
vehicles <- read.csv(unz("data/vehicles.csv.zip", "vehicles.csv"), stringsAsFactors = F)

#view the first few lines in the data set
head(vehicles)

#load variables information
labels <- do.call(rbind, strsplit(readLines("data/variables.txt"), " - "))

#view the first few lines of variables
head(labels)
```

#Exploring and describing fuel efficiency
Get some basic information on what the data set looks like. Some basic analysis will also be done.
```{r}
#Number of rows (observations) in the data set
nrow(vehicles)

#number of columns (variables) in the data set
ncol(vehicles)

#varible names (column headers)
names(vehicles)

#get the number of unique years
length(unique(vehicles[,"year"]))

#the first year (min function)
first_year <- min(vehicles[, "year"])
first_year

last_year <- max(vehicles[, "year"])
last_year

#table of fuel types and counts in the data set
table(vehicles$fuelType1)

#exploring transmission types
#set missing data to NA
vehicles$trany[vehicles$trany == ""] <- NA

#create a new variable trany2
#set variable to Auto if first four characters to autao and manual if not
vehicles$trany2 <- ifelse(substr(vehicles$trany, 1, 4) == "Auto", "Auto", "Manual")

#Convert the new variable to a factor and use table to see the distribution of values
vehicles$trany2 <- as.factor(vehicles$trany2)
table(vehicles$trany2)
```

#Analyzing automobile fuel efficiency over time
THis section analyzes fuel efficiency over time in relation to other data points.

##Graphing fuel efficiency by year for all cars
```{r}
#create a variable mpgByYr this is miles per gallon by year
mpgByYr <- ddply(vehicles, ~year, summarise, avgMPG = mean(comb08), avgHghy = mean(highway08), avgCity = mean(city08))

ggplot(mpgByYr, aes(year, avgMPG)) + geom_point() + geom_smooth() + xlab("Year") + ylab("Average MPG") + ggtitle("All cars")
```

It appears that fuel economy has increased a good amount since 2010. However, that could be because there has been an increase in other types of fuel that is used by other cars (electric, hybrid)
```{r}
table(vehicles$fuelType1)
```

##Gasoline cars only
This graph will filter out just cars that run off gasolineShow in New WindowClear OutputExpand/Collapse Output
```{r}
gasCars <- subset(vehicles, fuelType1 %in% c("Regular Gasoline", "Premium Gasoline", "Midgrade Gasoline") & fuelType2 == "" & atvType != "Hybrid")

mpgByYr_Gas <- ddply(gasCars, ~year, summarise, avgMPG = mean(comb08))

ggplot(mpgByYr_Gas, aes(year, avgMPG)) + geom_point() + geom_smooth() + xlab("Year") + ylab("Average MPG") + ggtitle("Gasoline cars")
```

##Graph fuel economy vs engine size
Looking to see if bigger engines have a lower fuel economy
```{r}
#Data type of displ
typeof(gasCars$displ)

#convert displ to a numeric variable
gasCars$displ <- as.numeric(gasCars$displ)

#create graph
ggplot(gasCars, aes(displ, comb08)) + geom_point() + geom_smooth()
```

##Year vs average enginge displacement (size)
Looking to see if more smaller engines were made in later years
```{r}
#Get average car size by year
avgCarSize <- ddply(gasCars, ~year, summarise, avgDispl = mean(displ))

#graph size vs year
ggplot(avgCarSize, aes(year, avgDispl)) + geom_point() + geom_smooth() + xlab("Year") + ylab("Average engine displacement (l)")
```

##Looking at average mpg vs average displacement in text form
```{r}
#Create a data frame with year average mpg and average displacement
byYear <- ddply(gasCars, ~year, summarise, avgMPG = mean(comb08), avgDispl = mean(displ))
head(byYear)

#Convert byYear to a long form showing year, variable, and value in each row
byYear2 = melt(byYear, id = "year")
levels(byYear2$variable) <- c("Average MPG", "Avg engine displacement")

head(byYear2)
```

##Create 2 graphs on top of each other
Create graphs that show average mpg on top and average displacement on bottom
```{r}
ggplot(byYear2, aes(year, value)) + geom_point() + geom_smooth() + facet_wrap(~variable, ncol = 1, scales = "free_y") + xlab("Year") + ylab("")
```

##Create a box and whisker plot of 4 cyliner engines seperated by auto and manual
In this section a graph is created that shows average mpg for 4 cylinder engines over time separated by auto and manual transaction
```{r}
#filter out 4 cylinder engines
gasCars4 <- subset(gasCars, cylinders == "4")

#build the graph
ggplot(gasCars4, aes(factor(year), comb08)) + geom_boxplot() + facet_wrap(~trany2, ncol = 1) + theme(axis.text.x = element_text(angle = 45)) + labs(x = "Year", y = "MPG")
```

##Proportion of manual vs automatic cars each year
This section will create a graph of proportions of automatic and manual transmissions each year
```{r}
ggplot(gasCars4, aes(factor(year), fill = factor(trany2))) + geom_bar(position = "fill") + labs(x = "Year", y = "Proprotion of cars", fill = "Transmission") + theme(axis.text.x = element_text(angle = 45)) + geom_hline(yintercept = 0.5, linetype = 2)
```
