---
title: "Assignment 6"
author: "Robert Evans"
date: "April 23, 2017"
output: word_document
---

#Chapter 7- Outlier Detection
This section goes over the various way of detecting outliers.

##7.1 Univariate Outlier Detection
This section goes over how to set up a univariate outlier detction. It then demonstrates how to apply it to multivariate data. This is achieved through the use of box and whisker plots that will show any outliers
```{r}
#set a seed for repeatable results
set.seed(3147)
#create an array of 100 random variables
x <- rnorm(100)
#print a summary
summary(x)

#detect outliers
boxplot.stats(x)$out

#plot the boxplot
boxplot(x)

#generate another 100 random numbers
y <- rnorm(100)
#combine x and y into a data frame
df <- data.frame(x, y)
#remove x and y to save memory
rm(x,y)
#print first 6 rows of data
head(df)

attach(df)
#get the index of outliers from x
(a <- which(x %in% boxplot.stats(x)$out))

#find the index of outliers from y
(b <- which(y %in% boxplot.stats(y)$out))

detatch(df)

# outliers in both x and y
(outlier.list1 <- intersect(a,b))

#plot the data frame and highlight outliers in red +
plot(df)
points(df[outlier.list1,], col="red", pch="+", cex=2)

#plot outliers in either x or y
(outlier.list2 <- union(a,b))

plot(df)
points(df[outlier.list2,], col="blue", pch="x", cex=1.5)

```

##7.2 Outlier Detection with LOF
Local Outlier Focus (LOF) is used for identifying density-based locaal outliers. In this method the density of a point is compared with its neighbors. If the density is significantly lower then it is considered an outlier.
```{r}
#Download and install package
#install.packages("DMwR")

#load required package
library(DMwR)

#remove the categorical column species
iris2 <- iris[,1:4]
#score data points for outlier detection
outlier.scores <- lofactor(iris2, k=5)
plot(density(outlier.scores))

#pick top five as outliers
outliers <- order(outlier.scores, decreasing=T)[1:5]
#print the outlier indexes
print(outliers)
#print outlier info
print(iris2[outliers,])

#show the outliers with a biplot of the two principal components
n <- nrow(iris2)
labels <- 1:n
labels[-outliers] <- "."
biplot(prcomp(iris2), cex=.8, xlabs=labels)

#plot outliers for all variables
pch <- rep(".", n)
pch[outliers] <- "+"
col <- rep("black", n)
col[outliers] <- "red"
pairs(iris2, pch=pch, col=col)

#install the Rlof package. This package allows for multiple number of neighbors
install.packages("Rlof")
library(Rlof)
outlier.scores <- lof(iris2, k=5)
#try with different # of neighbors (5-10)
outlier.scores <- lof(iris2, k=c(5:10))
```

##7.3 Outlier Detection by Clustering
This section uses the DBSCAN method of clustering for outlier detection. In this method data points are grouped into clusters if they are connected by a densely populated area. If a data point does not get grouped it is consider an outlier.
```{r}
#remove species from the data set
iris2 <- iris[,1:4]
#create the clusters
kmeans.result <- kmeans(iris2, centers=3)
#display cluster centers
kmeans.result$centers

#display cluster IDs
kmeans.result$cluster

#calculate distances between objects and centers
centers <- kmeans.result$centers[kmeans.result$cluster, ]
distances <- sqrt(rowSums((iris2 - centers)^2))

#pick top 5 largest distances
outliers <- order(distances, decreasing=T)[1:5]

#display outlier ids
print(outliers)

#display outlier information
print(iris2[outliers,])

#plot the data with outliers and cluster centers
plot(iris2[,c("Sepal.Length", "Sepal.Width")], pch="o",
     col=kmeans.result$cluster, cex=0.3)
#plot the center of cluster
points(kmeans.result$centers[,c("Sepal.Length","Sepal.Width")], col=1:3,
       pch=8, cex=1)
#plot the outliers
points(iris2[outliers, c("Sepal.Length", "Sepal.Width")], pch="+", col=4)
```

