---
title: "Extra Credit Assignment 3"
author: "Robert Evans"
date: "April 23, 2017"
output: word_document
---

#Chapter 4- Modeling Stock Market Data

##Introduction
Required packages are downloaded, installed, and loaded. Among these packages are:

* XML
* ggplot2
* plyr
* reshape2
* zoo
```{r}
#install.packages("XML")
#install.packages("gglot2")
#install.packages("plyr")
#install.packages("reshape2")
#install.packages("zoo")

library(XML)
library(ggplot2)
library(plyr)
library(reshape2)
library(zoo)
```

##Acquiring stock market data
In this section the data is downloaded from the finviz.com website. However, since the stock data is already provided it will not be downloaded.
```{r}
#build download url.
#url_to_open <- sprintf("http://finviz.com/export.ashx?v=152&c=%s", paste(0:68, collapse = ","))

#download csv file
#finviz <- read.csv(url(url_to_open))

#load the csv file
finviz <- read.csv("data/finviz.csv")
```

##Summarizing the data
Look at the first few records and a summary of the data
```{r}
#first 6 records
head(finviz)

#summary of each attribute
summary(finviz)
```

##Cleaning and exploring the data
In this section the data is cleaned (removal of percentage signs, dollar signs, commas, etc). Then some summaries of the data are run
```{r}
#create function to clean the data
clean_numeric <- function(s){
  s <- gsub("%|\\$|,|\\)|\\(", "", s)
  s <- as.numeric(s)
}

#clean the data
finviz <- cbind(finviz[,1:6],apply(finviz[,7:68], 2, clean_numeric))

#build a histogram of prices. One thing to note is with extremely high stock prices this will skew the graph making it useless.
hist(finviz$Price, breaks=100, main="Price Distribution", xlab="Price")

#put a cap of $150/share to get a useable graph
hist(finviz$Price[finviz$Price<150], breaks=100, main="Price Distribution", xlab="Price")

#Create a graph of sector specific prices.
#In the previous graph the majority of stocks were priced under $50. This means everything over that would be considered expensive. However, in some sectors the sector may have an average of $200/share which means a stock price of $100/share would be cheap.

#aggregate the sectors by average
sector_avg_prices <- aggregate(Price~Sector, data=finviz, FUN="mean")
colnames(sector_avg_prices)[2] <- "Sector_Avg_Price"

#Create the graph
ggplot(sector_avg_prices, aes(x=Sector, y=Sector_Avg_Price, fill=Sector)) +
  geom_bar(stat="identity") + ggtitle("Sector Avg Prices") +
  theme(axis.text.x = element_text(angle = 90, hjust=1))

#Looking into the financial sector

#Create a summary of the average price by industry
industry_avg_prices <- aggregate(Price~Sector+Industry, data=finviz,FUN="mean")
industry_avg_prices <- industry_avg_prices[order(industry_avg_prices$Sector,industry_avg_prices$Industry),]
colnames(industry_avg_prices)[3] <- "Industry_Avg_Price"

#Isolate the conglomerates sector
industry_chart <- subset(industry_avg_prices,Sector=="Financial")

#create the chart
ggplot(industry_chart, aes(x=Industry, y=Industry_Avg_Price, fill=Industry)) +
  geom_bar(stat="identity") + theme(legend.position="none") + ggtitle("Industry Avg Prices") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#It appears that Property & Casualty Insurance is the high industry. We will drill down into that to determine which company is causing the high stock price
company_chart <- subset(finviz, Industry=="Property & Casualty Insurance")

#create the chart
ggplot(company_chart, aes(x=Company, y=Price, fill=Company)) +
  geom_bar(stat="identity") + theme(legend.position="none") +
  ggtitle("Company Avg Prices") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#It appears that the company with the high stock price is Berkshire Hathaway with a stock price of over $400K/share. We will remove that to get a more realistic average
#Note the book shows only BRK-A. In this data set there is BRK-A and BRK-B. I will remove both of those. #In the conglomerates category I will also remove SEB to get a more realistic chart in the end as well.

finviz <- subset(finviz, Ticker!="BRK-A")
finviz <- subset(finviz, Ticker!="BRK-B")
finviz <- subset(finviz, Ticker!="SEB")

#recreate sector averages
sector_avg_prices <- aggregate(Price~Sector, data=finviz, FUN="mean")
colnames(sector_avg_prices)[2] <- "Sector_Avg_Price"

#create the graph
ggplot(sector_avg_prices, aes(x=Sector, y=Sector_Avg_Price, fill=Sector)) +
  geom_bar(stat="identity") + ggtitle("Sector Avg Prices") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

##Generating relative valuations
In this section a valuation of the stocks is made. This will run a relative valuation where the stock price is compared against different ratios to arrive at a valuation of the stock. This is easier than an intrinsic valuation because it does not require digging into the companies financials.
```{r}
#The first step is to melt the data. This converts the data from a wide format to a long one. This allows us to create an average of the different attributes grouped by sector. In the end the data will go back to a wide form with summaries
sector_avg <- melt(finviz, id="Sector")

#Filter the data out to just the fields that we want averages for
sector_avg <- subset(sector_avg, variable%in%c("Price", "P.E", "PEG", "P.S", "P.B"))

#Remove Missing values and average the values
sector_avg <- (na.omit(sector_avg))
sector_avg$value <- as.numeric(sector_avg$value)

#convert the data back to a wide form to create a summary by sector
sector_avg <- dcast(sector_avg, Sector~variable, mean)
colnames(sector_avg)[2:6] <- c("SAvgPE", "SAvgPEG", "SAvgPS", "SAvgPB", "SAvgPrice")

#do the same thing but at the industry level
industry_avg <- melt(finviz, id=c("Sector","Industry"))
industry_avg <- subset(industry_avg,variable %in% c("Price","P.E","PEG","P.S","P.B"))
industry_avg <- (na.omit(industry_avg))
industry_avg$value <- as.numeric(industry_avg$value)
industry_avg <- dcast(industry_avg, Sector+Industry~variable, mean)
industry_avg <- (na.omit(industry_avg))
colnames(industry_avg)[3:7] <- c("IAvgPE","IAvgPEG","IAvgPS","IAvgPB","IAvgPrice")

#Add the sector and industry averages to the original data set
finviz <- merge(finviz, sector_avg, by.x="Sector", by.y="Sector")
finviz <- merge(finviz, industry_avg, by.x=c("Sector","Industry"), by.y=c("Sector","Industry"))

#add field to determine if stock is undervalued (1) or overvalued (0) if the stock value is less than the average
finviz$SPEUnder <- 0
finviz$SPEGUnder <- 0
finviz$SPSUnder <- 0
finviz$SPBUnder <- 0
finviz$SPriceUnder <- 0
finviz$IPEUnder <- 0
finviz$IPEGUnder <- 0
finviz$IPSUnder <- 0
finviz$IPBUnder <- 0
finviz$IPriceUnder <- 0

#Create the valuations. Replace the 0 with 1 if the value is less than the average
finviz$SPEUnder[finviz$P.E<finviz$SAvgPE] <- 1
finviz$SPEGUnder[finviz$PEG<finviz$SAvgPEG] <- 1
finviz$SPSUnder[finviz$P.S<finviz$SAvgPS] <- 1
finviz$SPBUnder[finviz$P.B<finviz$SAvgPB] <- 1
finviz$SPriceUnder[finviz$Price<finviz$SAvgPrice] <- 1
finviz$IPEUnder[finviz$P.E<finviz$IAvgPE] <- 1
finviz$IPEGUnder[finviz$PEG<finviz$IAvgPEG] <- 1
finviz$IPSUnder[finviz$P.S<finviz$IAvgPS] <- 1
finviz$IPBUnder[finviz$P.B<finviz$IAvgPB] <- 1
finviz$IPriceUnder[finviz$Price<finviz$IAvgPrice] <- 1

#create a valuation index based on the sum of the above valuations. Higher numbers in this column represent undervalued stocks
finviz$RelValIndex <- apply(finviz[79:88],1,sum)
```

