---
title: "Assignment 3"
author: "Robert Evans"
date: "April 2, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#8.5.1- Classification with Original Data
The ctree() package is used to classify time series data with original data. In this section the class labels get changed into categorical values before the classification takes place to prevent class lables being interpreted as real numbers.
```{r}
sc <- read.table("data/synthetic_control.data", header=F, sep="")
classId <- c(rep("1",100), rep("2",100), rep("3",100),
             rep("4",100), rep("5",100), rep("6",100))
newSc <- data.frame(cbind(classId, sc))

library(party)
ct <- ctree(classId ~ ., data=newSc,
            controls = ctree_control(minsplit=30, minbucket=10, maxdepth =5))
pClassId <- predict(ct)
table(classId, pClassId)

#accuracy
(sum(classId==pClassId))/nrow(sc)

#plot decision tree
plot(ct, ip_args=list(pval=FALSE), ep_args=list(digits=0))
```

#8.5.2- Classification with Extracted Features
Extract features using a discrete wavelet transform (DWT)
```{r}
#install.packages("wavelets")
library(wavelets)
wtData <- NULL
for(i in 1:nrow(sc)) {
  a <- t(sc[i,])
  wt <- dwt(a, filter="haar", boundary="periodic")
  wtData <- rbind(wtData,
                  unlist(c(wt@W, wt@V[[wt@level]])))
}
wtData <- as.data.frame(wtData)
wtSc <- data.frame(cbind(classId, wtData))

#Create a decision tree with dwt
ct <- ctree(classId ~ ., data = wtSc,
            controls = ctree_control(minsplit=30, minbucket=10, maxdepth=5))
pClassId <- predict(ct)
table(classId, pClassId)

#accuarcy rate
(sum(classId==pClassId)) / nrow(wtSc)

#plot the decision tree
plot(ct, ip_args=list(pval=FALSE), ep_args=list(digits=0))
```

#8.5.3- k-NN Classification
Create a model of k-NN classification to find the k nearest neighbors of a new instance. The DTW method is an invalid distance method. This block does not run. Commented out so that I can get it to compile to a word document
```{r}
#k <- 20
#newTs <- sc[501,] + runif(100)*15
#distances <- dist(newTS, sc, method="DTW")
#s <- sort(as.vector(distances), index.return=TRUE)
# class IDs of k nearest neighbors
#table(classId[s$ix[1:k]])
```

