---
title: "Extra Credit Assignment 2"
author: "Robert Evans"
date: "April 17, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Simulating American Football Data
This assignment will download and process football data. After doing so games will be simulated

##Introduction
Install and load required packages
```{r}
#install.packages("XML")
#install.packages("RSQLite")
#install.packages("stringr")
#install.packages("ggplot2")

library(XML)
library(RSQLite)
library(stringr)
library(ggplot2)
```

##Acquiring and cleaning football data
In this section data from yahoo sports will be downloaded and cleaned. As a note the year will be able to be changed simply by changing the year variable
```{r}
#Set the year
year <- 2013

#download offensive data
url <- paste("http://sports.yahoo.com/nfl/stats/byteam?group=Offense&cat=Total&conference=NFL&year=season_",year,"&sort=530&old_category=Total&old_group=Offense")

#process offensive data
offense <- readHTMLTable(url, encoding = "UTF-8", colClasses="character")[[7]]

#clean the offensive data. Remove blank columns and set data types of other columns
offense <- offense[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28)]
offense[,1] <- as.character(offense[,1])
offense[,2:13] <- apply(offense[,2:13],2,as.numeric)
offense[,14] <- as.numeric(substr(offense[,14], 1, 2)) * 60 + as.numeric(substr(offense[,14], 4, 6))

#download defensive data
url <- paste("http://sports.yahoo.com/nfl/stats/byteam?group=Defense&cat=Total&conference=NFL&year=season_",year,"&sort=530&old_category=Total&old_group=Defense")

#process offensive data
defense <- readHTMLTable(url, encoding = "UTF-8", colClasses="character")[[7]]

#clean the defensive data. Remove blank columns and set data types of other columns
defense <- defense[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28)]
defense[,1] <- as.character(defense[,1])
defense[,2:13] <- apply(defense[,2:13],2,as.numeric)
```

##Analyzing and understanding football data
Run some basic analysis of the data and create some graphs to try and gain some useful insight into the data.
```{r}
#combine the offense and defense data frames to make it easier to explore
combined <- merge(offense, defense, by.x="Team", by.y="Team")

#reanme columns that have the same name
colnames(combined)[2] <- "Games"
colnames(combined)[3] <- "OffPPG"
colnames(combined)[4] <- "OffYPG"
colnames(combined)[5] <- "OffPassYPG"
colnames(combined)[6] <- "OffRushYPG"
combined$G.y <- NULL
colnames(combined)[15] <- "DefPPG"
colnames(combined)[16] <- "DefYPG"
colnames(combined)[17] <- "DefRushYPG"
colnames(combined)[18] <- "DefPassYPG"

#create a historam of offensive points per game. Points along the x-axis and number of teams on the y-axis.
hist(combined$OffPPG, breaks=10, main="Offensive Points Per Game", xlab="Offensive PPG", ylab="Number of Teams")

#show some basic statistics on offensive points per game
mean(combined$OffPPG)
sd(combined$OffPPG)
max(combined$OffPPG)
min(combined$OffPPG)

#create a historam of defensive points per game. Points along the x-axis and number of teams on the y-axis.
hist(combined$DefPPG, breaks=10, main="Defensive Points Per Game", xlab="Defensive PPG", ylab="Number of Teams")

#create a historam of first downs per game. First Downs Per Game along the x-axis and number of teams on the y-axis.
hist(combined$"1stD/G", breaks=10, main="Offensive 1st Downs Per Game", xlab="1st Downs/Game", ylab="Number of Teams")

#Create a bar chart of average points per grame grouped by team
ppg <- transform(combined, Team=reorder(Team,combined$OffPPG))
ggplot(ppg,aes(x=Team, y=OffPPG)) +
  geom_bar(stat='identity', color="black", fill="Blue") +
  coord_flip() + labs(x="Team", y="Avg Points Per Game") +
  ggtitle("Avg Points per Game") + theme(plot.title= element_text(size=18, face="bold"))

#Create a bar chart of ddefensive yards allowed per game grouped by team
ypg <- transform(combined, Team=reorder(Team,combined$DefYPG))
ggplot(ypg,aes(x=Team, y=DefYPG)) +
  geom_bar(stat='identity', color="black", fill="blue") +
  coord_flip() + labs(x="Team", y="Avg Yards Allowed per Game") +
  ggtitle("Avg Yards Allowed per Game") + theme(plot.title = element_text(size = 18, face = "bold"))

#Create a scatter plot to show the correlation between Offensive yards per game vs. offensive points per game
ggplot(combined, aes(x=combined$OffYPG, y=combined$OffPPG)) +
  geom_point(ize=2) + geom_smooth() +
  labs(x="Yards per Game", y="Points per Game") +
  ggtitle("Offense Yards vs. Points per Game") +
  theme(plot.title = element_text(size=18, face="bold"))

#Create a scatter plot to show the correlation between Deffensive Yards Per game vs. defensive points per game
ggplot(combined, aes(x=combined$DefYPG, y=combined$DefPPG)) +
  geom_point(shape=5, size=2) + geom_smooth() +
  labs(x="Yards Allowed per Game", y="Points Allowed Per Game") +
  ggtitle("Defense Yards vs. Points per Game") +
  theme(plot.title = element_text(size=18, face="bold"))

#create a scatter plot to show the correlation between time of posession and points per game
ggplot(combined, aes(x=combined$TOP, y=combined$OffPPG)) +
  geom_point(shape=5, size=2) + geom_smooth() +
  labs(x="Time of Possession (Seconds)", y="Points per Game") +
  ggtitle("Time of Possession vs. Points per Game") +
  theme(plot.title = element_text(size=18, face="bold"))
```

##Constructing indexes to measure offensive and defensive strength
In this section multiple fields are combined to create an index. These indexes measure how strong each team is on some aspect. They will be scaled from 0 to 100. 0 being the worst and 100 being the best.
```{r}
#create offensive massing strength index
offense$OPassStrength <- max(offense[,5])-offense[,5]
offense$OPassStrength <- (1-(offense$OPassStrength/max(offense$OPassStrength))) * 100

#create offensive rushing strength index
offense$ORushStrength <- max(offense[,6]) - offense[,6]
offense$ORushStrength <- (1-(offense$ORushStrength/max(offense$ORushStrength))) * 100

#offenseve points per game
offense$OPPGStrength <- max(offense[,3]) - offense[,3]
offense$OPPGStrength <- (1-(offense$OPPGStrength/max(offense$OPPGStrength))) * 100

#offense yards per game
offense$OYPGStrength <- max(offense[,4])-offense[,4]
offense$OYPGStrength <- (1-(offense$OYPGStrength/max(offense$OYPGStrength))) * 100

offense$OffStrength <- (offense$OPassStrength + offense$ORushStrength + offense$OPPGStrength + offense$OYPGStrength)/4

#create defensive massing strength index
defense$DPassStrength <- max(defense[,6])-defense[,6]
defense$DPassStrength <- defense$DPassStrength/max(defense$DPassStrength) * 100

#create defensive rushing strength index
defense$DRushStrength <- max(defense[,5]) - defense[,5]
defense$DRushStrength <- defense$DRushStrength/max(defense$DRushStrength) * 100

#defenseve points per game
defense$DPPGStrength <- max(defense[,3]) - defense[,3]
defense$DPPGStrength <- defense$DPPGStrength/max(defense$DPPGStrength) * 100

#defense yards per game
defense$DYPGStrength <- max(defense[,4])-defense[,4]
defense$DYPGStrength <- defense$DYPGStrength/max(defense$DYPGStrength) * 100

defense$DefStrength <- (defense$DPassStrength + defense$DRushStrength + defense$DPPGStrength + defense$DYPGStrength)/4
```

##Simulating a single game with outcomes decided by calculations
This section utilizes the indexes in order to determine what team SHOULD win a given game
```{r}
#Define teams
home_team <- "Chicago Bears"
away_team <- "New Orleans Saints"

#bring offense and defense strengths together
off_game <- subset(offense, Team==home_team | Team==away_team)[,c(1,15,16,19)]
def_game <- subset(defense, Team==home_team | Team==away_team)[,c(1,14,15,18)]

#merge the two frames together
game <- merge(off_game,def_game,by.x="Team",by.y="Team")

#Calculate net passing strength
game$Net_Pass[game$Team==home_team] <- game$OPassStrength[game$Team==home_team] - game$DPassStrength[game$Team==away_team]
game$Net_Pass[game$Team==away_team] <- game$OPassStrength[game$Team==away_team] - game$DPassStrength[game$Team==home_team]

#calculate net rushing strength
game$Net_Rush[game$Team==home_team] <- game$ORushStrength[game$Team==home_team] - game$DRushStrength[game$Team==away_team]
game$Net_Rush[game$Team==away_team] <- game$ORushStrength[game$Team==away_team] - game$DRushStrength[game$Team==home_team]

#determine the winner
#if the net total is greater for home team they win
game$Net_Total <- game$Net_Pass + game$Net_Rush
game$Net_Total

if(game$Net_Total[game$Team == home_team] >= game$Net_Total[game$Team==away_team]){
  winner <- home_team
  loser <- away_team
} else {
  winner <- away_team
  loser <- home_team
}
print(paste(winner, "beat", loser))
```

##Simulating multiple games with outcomes decided by calculations
This section does the same as the previous section except it will run for multiple teams
```{r}
#build the schedule of games
games_per_team <- 50

for(week in 1:games_per_team){
  home_index <- sample(1:32,16,replace=F)
  home_teams <- data.frame(HomeTeam=offense[home_index, 1])
  away_teams <- data.frame(AwayTeam=offense[-home_index, 1])
  
  if(week ==1){
    schedule <- cbind(Week=week,HomeTeam=home_teams,AwayTeam=away_teams)
  } else {
    temp <- cbind(Week=week,HomeTeam=home_teams,AwayTeam=away_teams)
    schedule <- rbind(schedule,temp)
  }
}

#build a record tracker
records <- data.frame(Team=offense$Team)
records$Wins <- 0
records$Losses <- 0

#Play the games
for(i in 1:nrow(schedule)){
  home_team <- schedule[i,2]
  away_team <- schedule[i,3]
  week <- schedule[i,1]
  
  #bring offense and defense strengths together
  off_game <- subset(offense, Team==home_team | Team==away_team)[,c(1,15,16,19)]
  def_game <- subset(defense, Team==home_team | Team==away_team)[,c(1,14,15,18)]

  #merge the two frames together
  game <- merge(off_game,def_game,by.x="Team",by.y="Team")
  
  #Calculate net passing strength
  game$Net_Pass[game$Team==home_team] <- game$OPassStrength[game$Team==home_team] - game$DPassStrength[game$Team==away_team]
  game$Net_Pass[game$Team==away_team] <- game$OPassStrength[game$Team==away_team] - game$DPassStrength[game$Team==home_team]
  
  #calculate net rushing strength
  game$Net_Rush[game$Team==home_team] <- game$ORushStrength[game$Team==home_team] - game$DRushStrength[game$Team==away_team]
  game$Net_Rush[game$Team==away_team] <- game$ORushStrength[game$Team==away_team] - game$DRushStrength[game$Team==home_team]
  
  #determine the winner
  #if the net total is greater for home team they win
  game$Net_Total <- game$Net_Pass + game$Net_Rush
  game$Net_Total

  if(game$Net_Total[game$Team == home_team] >= game$Net_Total[game$Team==away_team]){
    winner <- home_team
    loser <- away_team
  } else {
    winner <- away_team
    loser <- home_team
  }
  
  #update win/loss table and print the winner/loser
  if(i==1){
    winnerdf <- data.frame(Winner=winner)
    loserdf <- data.frame(Loser=loser)
    results <- cbind(winnerdf,loserdf)
  } else {
    winnerdf <- data.frame(Winner=winner)
    loserdf <- data.frame(Loser=loser)
    temp <- cbind(winnerdf,loserdf)
    results <- rbind(results,temp)
  }
  
  records$Wins[records$Team==winner] <- as.numeric(records$Wins[records$Team==winner]) + 1
  records$Losses[records$Team==loser] <- as.numeric(records$Losses[records$Team==loser]) + 1
  
  print(paste("Week", week,":", winner, "beat", loser))
}

records <- records[order(-records$Wins),]
records
```




