---
title: "Extra Credit Assignment 2"
author: "Robert Evans"
date: "April 17, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Simulating American Football Data
This assignment will download and process football data. After doing so games will be simulated

##Introduction
Install and load required packages
```{r}
#install.packages("XML")
#install.packages("RSQLite")
#install.packages("stringr")
#install.packages("ggplot2")

library(XML)
library(RSQLite)
library(stringr)
library(ggplot2)
```

##Acquiring and cleaning football data
In this section data from yahoo sports will be downloaded and cleaned. As a note the year will be able to be changed simply by changing the year variable
```{r}
#Set the year
year <- 2013

#download offensive data
url <- paste("http://sports.yahoo.com/nfl/stats/byteam?group=Offense&cat=Total&conference=NFL&year=season_",year,"&sort=530&old_category=Total&old_group=Offense")

#process offensive data
offense <- readHTMLTable(url, encoding = "UTF-8", colClasses="character")[[7]]

#clean the offensive data. Remove blank columns and set data types of other columns
offense <- offense[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28)]
offense[,1] <- as.character(offense[,1])
offense[,2:13] <- apply(offense[,2:13],2,as.numeric)
offense[,14] <- as.numeric(substr(offense[,14], 1, 2)) * 60 + as.numeric(substr(offense[,14], 4, 6))

#download defensive data
url <- paste("http://sports.yahoo.com/nfl/stats/byteam?group=Defense&cat=Total&conference=NFL&year=season_",year,"&sort=530&old_category=Total&old_group=Defense")

#process offensive data
defense <- readHTMLTable(url, encoding = "UTF-8", colClasses="character")[[7]]

#clean the defensive data. Remove blank columns and set data types of other columns
defense <- defense[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28)]
defense[,1] <- as.character(defense[,1])
defense[,2:13] <- apply(defense[,2:13],2,as.numeric)
```

##Analyzing and understanding football data
Run some basic analysis of the data and create some graphs to try and gain some useful insight into the data.
```{r}
#combine the offense and defense data frames to make it easier to explore
combined <- merge(offense, defense, by.x="Team", by.y="Team")

#reanme columns that have the same name
colnames(combined)[2] <- "Games"
colnames(combined)[3] <- "OffPPG"
colnames(combined)[4] <- "OffYPG"
colnames(combined)[5] <- "OffPassYPG"
colnames(combined)[6] <- "OffRushYPG"
combined$G.y <- NULL
colnames(combined)[15] <- "DefPPG"
colnames(combined)[16] <- "DefYPG"
colnames(combined)[17] <- "DefRushYPG"
colnames(combined)[18] <- "DefPassYPG"

#create a historam of offensive points per game. Points along the x-axis and number of teams on the y-axis.
hist(combined$OffPPG, breaks=10, main="Offensive Points Per Game", xlab="Offensive PPG", ylab="Number of Teams")

#show some basic statistics on offensive points per game
mean(combined$OffPPG)
sd(combined$OffPPG)
max(combined$OffPPG)
min(combined$OffPPG)

#create a historam of defensive points per game. Points along the x-axis and number of teams on the y-axis.
hist(combined$DefPPG, breaks=10, main="Defensive Points Per Game", xlab="Defensive PPG", ylab="Number of Teams")

#create a historam of first downs per game. First Downs Per Game along the x-axis and number of teams on the y-axis.
hist(combined$"1stD/G", breaks=10, main="Offensive 1st Downs Per Game", xlab="1st Downs/Game", ylab="Number of Teams")

#Create a bar chart of average points per grame grouped by team
ppg <- transform(combined, Team=reorder(Team,combined$OffPPG))
ggplot(ppg,aes(x=Team, y=OffPPG)) +
  geom_bar(stat='identity', color="black", fill="Blue") +
  coord_flip() + labs(x="Team", y="Avg Points Per Game") +
  ggtitle("Avg Points per Game") + theme(plot.title= element_text(size=18, face="bold"))

#Create a bar chart of ddefensive yards allowed per game grouped by team
ypg <- transform(combined, Team=reorder(Team,combined$DefYPG))
ggplot(ypg,aes(x=Team, y=DefYPG)) +
  geom_bar(stat='identity', color="black", fill="blue") +
  coord_flip() + labs(x="Team", y="Avg Yards Allowed per Game") +
  ggtitle("Avg Yards Allowed per Game") + theme(plot.title = element_text(size = 18, face = "bold"))

#Create a scatter plot to show the correlation between Offensive yards per game vs. offensive points per game
ggplot(combined, aes(x=combined$OffYPG, y=combined$OffPPG)) +
  geom_point(ize=2) + geom_smooth() +
  labs(x="Yards per Game", y="Points per Game") +
  ggtitle("Offense Yards vs. Points per Game") +
  theme(plot.title = element_text(size=18, face="bold"))

#Create a scatter plot to show the correlation between Deffensive Yards Per game vs. defensive points per game
ggplot(combined, aes(x=combined$DefYPG, y=combined$DefPPG)) +
  geom_point(shape=5, size=2) + geom_smooth() +
  labs(x="Yards Allowed per Game", y="Points Allowed Per Game") +
  ggtitle("Defense Yards vs. Points per Game") +
  theme(plot.title = element_text(size=18, face="bold"))

#create a scatter plot to show the correlation between time of posession and points per game
ggplot(combined, aes(x=combined$TOP, y=combined$OffPPG)) +
  geom_point(shape=5, size=2) + geom_smooth() +
  labs(x="Time of Possession (Seconds)", y="Points per Game") +
  ggtitle("Time of Possession vs. Points per Game") +
  theme(plot.title = element_text(size=18, face="bold"))
```

