---
title: "Assignment 1"
author: "Robert Evans"
date: "March 19, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Chapter 1
##Iris Data Set
```{r}
str(iris)
```

##Bodyfat Dataset
```{r}
#Commenting the install.packages out to prevent installation from running multiple times
#install.packages("mboost")
library(mboost)
#This no longer works data is now in TH.data
#source: http://www.rdatamining.com/books/rdm/faq/whereistofindthebodyfatdata
#data("bodyfat", package="mboost")
data("bodyfat", package = "TH.data")
str(bodyfat)
```

#Chapter 2- Data Import/Export

##2.1- Save/Load R Data
This will create an array of the numbers 1-10 and save it to data/dumData.Rdata. It will then delete the array "a" and reload the file and print out the file contents
```{r}
a <- 1:10
save(a, file="data/dumData.Rdata")
rm(a)
load("data/dumData.Rdata")
print(a)
```

##2.2- Import fram and Export to .CSV files
Here we create a data frame and save it to a .csv file and then reload the same .csv and print out the contents of it.
```{r}
var1 <- 1:5
var2 <- (1:5) / 10
var3 <- c("R", "and", "Data Mining", "Examples", "Case Studies")
a <- data.frame(var1, var2, var3)
names(a) <- c("VariableInt", "VariableReal", "VariableChar")
write.csv(a, "data/dummyData.csv", row.names = FALSE)
b <- read.csv("data/dummyData.csv")
print(b)
```

##2.3- Import Data From SAS
This section goes over how to import data from SAS. However, I am unable to do this as I do not have SAS installed on my system.

##2.4- Import/Export via ODBC
This ection goes over how to import data from a data base connection. I am unable to do this as I do not have an available database to connect to.

#Chapter 3- Data Exploration

##3.1- Have a Look at Data
This section discusses how to explore data with different methods. The IRIS dataset is used here.
```{r}
#Retrieves the dimensions of the data set
dim(iris)
#Gets the column names of the data
names(iris)
#Gets the structure of the data
str(iris)
#Gets the attributes
attributes(iris)
#Get the first 5 rows of data
#Notice that in R the array starts at 1 and not 0 like most other programming languages
iris[1:5,]
#Get the first part of the iris data set
head(iris)
#Get the last part of the iris data set
tail(iris)
#Retrieves the Sepal.Length of the first 10 rows
iris[1:10, "Sepal.Length"]
#This does the same as the last command
iris$Sepal.Length[1:10]
```

##3.2- Explore Individual Variables
This section covers how to explore the individual variables of a dataset

Summary calcuates the min, maax, meadian, the first and third quartiles, for a factor it shows counts of every level
```{r}
summary(iris)
```

Calculating the quantiles. It defaults to every 25%. However the quantiles can be changed.
```{r}
quantile(iris$Sepal.Length)
quantile(iris$Sepal.Length, c(.1, .3, .65))
```

Building a frequency table. Then generating a piechart off of that table
```{r}
table(iris$Species)
pie(table(iris$Species))
var(iris$Sepal.Length)
```

Generating a histogram and density plot of the Sepal.Length
```{r}
hist(iris$Sepal.Length)
plot(density(iris$Sepal.Length))
```

##3.3- Explore Multiple Variables

Calculating the covariance and correlation of two variables. This also calculates the distribution in subsets via aggregate
```{r}
cov(iris$Sepal.Length, iris$Petal.Length)
cor(iris$Sepal.Length, iris$Petal.Length)
cor(iris[,1:4])
aggregate(Sepal.Length ~ Species, summary, data=iris)
```

This will build a box plot. The box shows the interquartile range (25%-75% observation) the median (line in the middle) and the outliers.
```{r}
boxplot(Sepal.Length~Species, data=iris)
```

Creating a scatter plot with the Sepal.Length as the x-axis, Sepal.Width as the y-axis, and color coded species
```{r}
with(iris, plot(Sepal.Length, Sepal.Width, col=Species, pch=as.numeric(Species)))

```
Creating a scatter plot with the Sepal.Length as the x-axis, Sepal.Width as the y-axis, and species. This does the same thing but adds noise to the data.
```{r}
plot(jitter(iris$Sepal.Length), jitter(iris$Sepal.Width))
```

Creating a pairs plot. This shows correlation between the different attributes in the data set.
```{r}
pairs(iris)
```

##3.4 More Explorations
This covers additional explorations with the additional packages

Utilizing the scatterplot 3d package to create a scatterplot of Petal.Width, Sepal.Length, and Sepal.Width
```{r}
#install.packages("scatterplot3d")
library(scatterplot3d)
scatterplot3d(iris$Petal.Width, iris$Sepal.Length, iris$Sepal.Width)
```

Creating a level plot of Sepal.Length and Sepal.Width
```{r}
#install.packages("lattice")
library(lattice)
print(levelplot(Petal.Width~Sepal.Length*Sepal.Width, iris, cuts=9,
                col.regions=grey.colors(10)[10:1]))
```

Creating a contour plot of the volcano data set
```{r}
filled.contour(volcano, color=terrain.colors, asp=1,
               plot.axes=contour(volcano, add=T))
```

Creating a 3D visualization of the volcano dataset
```{r}
persp(volcano, theta = 25, phi = 30, expand = 0.5, col = "lightblue")
```

Creating an interactive 3D Scatter plot with the rgl package. This will not work in the output document so it has been commented out. It does create an interactive 3d scatter plot
```{r}
#install.packages("rgl")
#library(rgl)
#plot3d(iris$Petal.Width, iris$Sepal.Length, iris$Sepal.Width)
```

Parallel Coordinates Graph. It was unable to find this function so has been commented out.
```{r}
#parcoord(iris[1:4], col=iris$Species)
```

Parallel Coordinates Graph seperated by species
```{r}
parallelplot(~iris[1:4] | Species, data=iris)
```

Creating a graph with the ggplot2 library
```{r}
#install.packages("ggplot2")
library(ggplot2)
qplot(Sepal.Length, Sepal.Width, data=iris, facets=Species ~.)
```

##3.5 Save Charts Into Files
Saving charts into a pdf file
```{r}
pdf("myPlot.pdf")
x <- 1:50
plot(x, log(x))
graphics.off()
```
Saving charts into a postscript file
```{r}
postscript("myPlotPostScript.ps")
x <- -20:20
plot(x, x^2)
graphics.off()
```




