---
title: "Assignment 1"
author: "Robert Evans"
date: "March 19, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Chapter 1
##Iris Data Set
```{r}
str(iris)
```

##Bodyfat Dataset
```{r}
#Commenting the install.packages out to prevent installation from running multiple times
#install.packages("mboost")
library(mboost)
#This no longer works data is now in TH.data
#source: http://www.rdatamining.com/books/rdm/faq/whereistofindthebodyfatdata
#data("bodyfat", package="mboost")
data("bodyfat", package = "TH.data")
str(bodyfat)
```

#Chapter 2- Data Import/Export

##2.1- Save/Load R Data
This will create an array of the numbers 1-10 and save it to data/dumData.Rdata. It will then delete the array "a" and reload the file and print out the file contents
```{r}
a <- 1:10
save(a, file="data/dumData.Rdata")
rm(a)
load("data/dumData.Rdata")
print(a)
```

##2.2- Import fram and Export to .CSV files
Here we create a data frame and save it to a .csv file and then reload the same .csv and print out the contents of it.
```{r}
var1 <- 1:5
var2 <- (1:5) / 10
var3 <- c("R", "and", "Data Mining", "Examples", "Case Studies")
a <- data.frame(var1, var2, var3)
names(a) <- c("VariableInt", "VariableReal", "VariableChar")
write.csv(a, "data/dummyData.csv", row.names = FALSE)
b <- read.csv("data/dummyData.csv")
print(b)
```

##2.3- Import Data From SAS
This section goes over how to import data from SAS. However, I am unable to do this as I do not have SAS installed on my system.

##2.4- Import/Export via ODBC
This ection goes over how to import data from a data base connection. I am unable to do this as I do not have an available database to connect to.